Model Checking

### 2023-05-29 owen/Write Ups/03_existing_models.qmd
add comments to make self contained

@@

@@

@@

This model gives similar predictive performance to the previously proposed
model.

>>>>

```{r process data for this model}
#| eval: false

lichess_data <- readRDS("../../rdata/lichess_pilot.RData")

lichess_data %>% 
  filter(Event == "Rated Bullet game") %>% 
  group_by(TimeControl) %>% 
  count() %>% 
  arrange(-n)

bullet_60 <- lichess_data %>% 
  filter(Event == "Rated Bullet game") %>% 
  filter(TimeControl == "60+0")

get_hist <- function(user, games, prev_n) {
  hist_games <- games %>% 
    filter(White == user | Black == user) %>% 
    arrange(UTCDate, UTCTime) %>% 
    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
    select(White:BlackElo, focal_white) %>% 
    mutate(focal_result = case_when(
      (focal_white == 1 & Result == "1-0") ~ 1,
      (focal_white == 0 & Result == "0-1") ~ 1,
      (Result == "1/2-1/2") ~ 0.5,
      .default = 0
      )) %>% 
    mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
                              roll_mean(focal_result, n = prev_n)))

  hist_games
}

focal_players <- head(sort(table(c(bullet_60$Username)), decreasing = TRUE), 18)

tidy_games <- map_dfr(top_players, get_hist, bullet_60, prev_n = 10) %>% 
  as_tibble()

init_data <- tidy_games %>% 
  mutate(WhiteElo = as.numeric(WhiteElo), 
         BlackElo = as.numeric(BlackElo)) %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
  mutate(elo_diff = ifelse(focal_white == 1, 
                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
  mutate(focal_id = match(focal_user, top_players)) %>% 
  select(focal_user, focal_id, focal_white, 
         focal_win_prop, elo_diff, focal_result) %>% 
  group_by(focal_id) %>% 
  mutate(ave_prop = lag(focal_win_prop, default = 0) - mean(focal_win_prop)) %>% 
  filter(focal_result != 0.5)

```{r fit stan model}
#| eval: false

## takes approx 3 minutes on m1 pro mac

stan_data_ave <- list(N = nrow(init_data),
                  J = length(top_players),
                  y = init_data$focal_result,
                  id = init_data$focal_id,
                  colour = init_data$focal_white,
                  elo = init_data$elo_diff,
                  win_prop = init_data$ave_prop)

stan_file <- "../model3.stan"

mod <- cmdstan_model(stan_file)

fit3_ave <- mod$sample(data = stan_data_ave,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4,
                  refresh = 100)

fit3_ave$save_object("../model_fits/init_model_n10.RDS")
```

@@

```{r data for model 2}
#| eval: false

hist_data_init <- tidy_games %>% 
  mutate(WhiteElo = as.numeric(WhiteElo), 
         BlackElo = as.numeric(BlackElo)) %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
  mutate(elo_diff = ifelse(focal_white == 1, 
                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
  mutate(focal_id = match(focal_user, top_players)) %>% 
  select(focal_user, focal_id, focal_white, 
         focal_win_prop, elo_diff, focal_result) %>% 
  group_by(focal_id) %>% 
  mutate(ave_prop = lag(focal_win_prop, n = 2, default = 0) - 
           mean(focal_win_prop),
         prev_game = lag(focal_result, default = 0)) %>% 
  filter(focal_result != 0.5)

stan_data_hist <- list(N = nrow(hist_data_init),
                  J = length(top_players),
                  y = hist_data_init$focal_result,
                  id = hist_data_init$focal_id,
                  colour = hist_data_init$focal_white,
                  elo = hist_data_init$elo_diff,
                  win_prop = hist_data_init$ave_prop,
                  prev_game = hist_data_init$prev_game)

```{r fit the stan model}
#| eval: false

## takes approx 4 minutes on m1 pro mac

stan_file <- "../model4.stan"

mod <- cmdstan_model(stan_file)

fit4 <- mod$sample(data = stan_data_hist,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4,
                  refresh = 100)

fit4$summary()

fit4$save_object(file = "../model_fits/model4.RDS")
```

@@

### this takes several hours, needs to be run on a cluster

@@

This model gives similar predictive performance to the previously proposed
model. However, there are some minor concerns with model fitting 
which should be addressed (convergence issues).
